Reproducible and programmatic generation of neuroimaging visualisations

### 2022-04-02 manuscript.Rmd
figure compiled

---
title: "Reproducible and programmatic generation of neuroimaging visualisations"
author: "Sidhant Chopra, Löic Labache, Winnie Orchard, [...]"
date: "09/03/2022"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

@@

### Figure 1

```{r Figure1 , echo=FALSE}
library(patchwork)
library(ggplot2)

#loop this?
fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1h', 'fig1i')
figs <- list()

@@

patchwork <- figs[[1]]/figs[[2]]|(figs[[3]]|figs[[6]])/figs[[4]]|(figs[[5]]|figs[[8]])/figs[[7]]|figs[[9]]/figs[[10]]

@@

\
*[Figure 1 - Add Caption]*

@@

[view_xifti_volume(xii, interactive = TRUE) in cifti tools]

@@

-   While most of these tools do not require a strond knowledge of
    programming, there can still be a steeper learning curve when
    compared to GUI-based tools.

-   Leged placement and creation

@@

-   But fsbrain has a function;
    <https://github.com/dfsp-spirit/fsbrain/blob/cb1b9d667a9449b214b20f6420038af2541e48b4/R/vis_dti_tracks.R>)

-   For tools which require non-standard or uncommon atlases or ROIs.

-   Although, with `ciftitools`, this becomes easier, as long as atlas
    follows the CIFTI convention.

-   3D renderings of non-standard ROIs like unique subcortical schemes
    or brain stem are still difficult to do.

-   Have to use complex tools like pyVista

@@

png("data/fig1a.png") 
ortho2(x = template, 
       y = effect, 
       crosshairs = F, 
       bg = "white", 
       NA.x = T, 
       col.y= hotmetal(), 
       xyz = c(70,50,80),
       useRaster = T,
       ycolorbar = TRUE, 
       mfrow = c(1,3)) +
  colorbar(breaks,col = hotmetal(n=length(breaks)-1), 
           text.col = "black", labels = TRUE, maxleft = 0.95) + 
  title(main = "Voxel Statistic Map")
dev.off()

@@

png('data/fig1b.png')
overlay(x=robust_window(dti), y=atlas, 
        plot.type = "single", 
        z=c(seq(30,110,10)), 
        col.y =  alpha(rainbow(300), 0.3),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, 
        zlim.y =c(1,300), 
        mar = c(0.1, 0.1, 1.5, 0.1)) + #marget of bottom, left, top, right borders
  title("Voxel Atlas Overlay", line = 1.3)
dev.off()

@@

view_xifti_surface(cii, 
                   hemisphere = "left", 
                   view = "both", 
                   idx = 1,
                   colors = viridis::viridis(n = 100),
                   zlim = c(1,1.5),  
                   legend_embed = T,
                   title = "Vertex Statistic Map", 
                   cex.title = 2)

@@

#fig1g <- cowplot::plot_grid(cortex, subcortex, nrow = 1, rel_widths = c(1,.7))

@@

>>>>

---
title: "Reproducible and programmatic generation of neuroimaging visualisations"
author: "Sidhant Chopra, Löic Labache, Winnie Orchard, [...]"
date: "09/03/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

@@

\newpage
\blandscape

### Figure 1 [Fix image size ratios]

```{r Figure1 , echo=FALSE, out.height = '80%', out.width = '140%'}
library(patchwork)
library(ggplot2)

#loop this?
fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
figs <- list()

@@

patchwork <- figs[[1]]/figs[[2]]|(figs[[3]]|figs[[6]])/figs[[4]]|(figs[[5]]|(figs[[8]]/figs[[9]]))/figs[[7]]|figs[[10]]/figs[[11]]

@@

\
*[Figure 1 - Add Caption and remove panel F]*

\elandscape

@@

@@

-   While most of these tools do not require a strong knowledge of
    programming, there can still be a steeper learning curve when
    compared to GUI-based tools.

@@

-   Lack of user friendly tools which can render non-standard or uncommon atlases or ROIs. Although, with `ciftitools`, this becomes easier, as long as atlas follows the CIFTI convention.

-   3D renderings of non-standard ROIs like unique subcortical schemes
    or brain stem are still difficult to do. Can to use complex workarounds like ITKsnap + pyVista.

### Supplement (Code used to generate figure)

@@

png("data/fig1a.png") 
ortho2(x = template, 
       y = effect, 
       crosshairs = F, 
       bg = "white", 
       NA.x = T, 
       col.y= hotmetal(), 
       xyz = c(70,50,80),
       useRaster = T,
       ycolorbar = TRUE, 
       mfrow = c(1,3)) +
  colorbar(breaks,col = hotmetal(n=length(breaks)-1), 
           text.col = "black", labels = TRUE, maxleft = 0.95) 
dev.off()

@@

png('data/fig1b.png')
overlay(x=robust_window(dti), y=atlas, 
        plot.type = "single", 
        z=c(seq(30,110,10)), 
        col.y =  alpha(rainbow(300), 0.3),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, 
        zlim.y =c(1,300))

      #  mar = c(0.1, 0.1, 1.5, 0.1)) + #marget of bottom, left, top, right borders
dev.off()

@@

view_xifti_surface(cii, 
                   hemisphere = "left", 
                   view = "both", 
                   idx = 1,
                   colors = viridis::viridis(n = 100),
                   zlim = c(1,1.5),  
                   legend_embed = T,
                   cex.title = 2)

@@

@@

```{r, eval=FALSE}

make_ggseg3d <- function(attribute, 
                         colour.pal = c("light yellow",
                                        "orange",
                                        "red",
                                        "dark red"), 
                         hide.colourbar=FALSE, 
                         output.png=FALSE, 
                         file.name="ggseg_3d.png") {
  # Inputs: 
  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left-Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Right-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Right-Amygdala, Right-Accumbens-area

  library(ggseg3d)
  library(tidyr)
  library(dplyr)
  #remove(aseg_3d)
  aseg_3d <- aseg_3d
  aseg_3d <- tidyr::unnest(aseg_3d, cols = c(ggseg_3d))

  attribute.ggseg3d <- c(rep(NA, 4), attribute[1:4], rep(NA, 3), attribute[5:7],
                         rep(NA, 5), attribute[8:14],  rep(NA, 6))

  data <- dplyr::mutate(aseg_3d, attribute =  attribute.ggseg3d)
  #remove NA regions
  aseg_3d[which(is.na(data$attribute)),] <- NA
  aseg_3d <- tidyr::drop_na(aseg_3d)

  data[which(is.na(data$attribute)),] <- NA
  data <- tidyr::drop_na(data)
  data$attribute[data$attribute==0]<-NA #make 0 valus NA so they are set as grey in ggseg3d

  scene=list(camera = list(eye = list(x = 0, y = 1, z = -2.25)),
             aspectratio = list(x=1.6,y=1.6,z=1.6))

  plot <- ggseg3d::ggseg3d(.data = data, 
                           atlas = aseg_3d, 
                           colour = "attribute", 
                           text = "attribute",
                           palette = colour.pal)
  plot <- remove_axes(plot)
  plot <- plotly::layout(plot, 
                         scene = scene, 
                         width = 600, height = 600)

  if(hide.colourbar==TRUE){
    plot <- plotly::hide_colorbar(plot)
  }

  if( output.png==TRUE){
    plotly::orca(plot, file =  file.name)
  }

  return(plot)

}

set.seed(1993)
volume <- sample(1:20, 14) #randomly generated data for 14 structures
make_ggseg3d(volume, 
             colour.pal = c("white", "firebrick"), 
             output.png = T, file.name = 
               "data/fig1g3.png")
